---
title: "dplyr::if_else()"
author: "Yang Zhu"
date: "9/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(lubridate)
```

We will discuss the difference between `base::ifelse()` and `dplyr::if_else`. I
would highly recommend to use dplyr::if\_else since it is more user friendly
when dealing with date objetcs and missing data.

ifelse(test, yes, no)
if_else(condition, true, false, missing = NULL)

## Dealing with missing data
```{r}
numeric_lst <- c(5, 4, 3, NA, NA, 2, 1, -1, 0)
ifelse(numeric_lst >= 3, "Big", "Small")
if_else(numeric_lst >= 3, "Big", "Small")
if_else(numeric_lst >= 3, "Big", "Small", missing = "Not Applicable")
```

## Coercion Problem

You have to keep that in mind. `ifelse()` would coerce the return values for
both TRUE and FALSE to same object type, but `if_else()` would not. It gives an
error message instead. Automatical coercion is not always good, you will see
examples later on.

```{r error=TRUE}
ifelse(numeric_lst >= 3, "Big", 0)
if_else(numeric_lst >= 3, "Big", 0)
```

## Replace with NA

Sometimes we do want to replace `NA` if condition is FALSE. However, you will
see error message when you are using `dplyr::if_else()`. Because there are
different kind of `NA` in R. In the case below, you should use `NA_character_`
instead.

```{r error=TRUE}
ifelse(numeric_lst >= 3, "Big", NA)
if_else(numeric_lst >= 3, "Big", NA)
if_else(numeric_lst >= 3, "Big", NA_character_)
```

## Dealing with date object

In the first two examples, you can see `base:ifelse()` function coerce all dates
into numeric object, but `dplyr::if_else()` does not. You should be very careful
with replace `NA` on date object, there is no `NA_date` in R. You could use
`as_date("")` as `NA` here.

```{r error=TRUE}
date_lst <- lubridate::as_date(c("2020-08-01", "2020-09-01", NA, "1999-01-01"))
# base:ifelse()
ifelse(date_lst > "2020-01-01", 0, date_lst)
wrong_output <- ifelse(date_lst > "2020-01-01", date_lst - ddays(1), date_lst); wrong_output
class(wrong_output)
# dplyr::if_else()
correct_output <- if_else(date_lst > "2020-01-01", date_lst - ddays(1), date_lst); correct_output
class(correct_output)
# replace with NA
if_else(date_lst > "2020-01-01", NA, date_lst)
if_else(date_lst > "2020-01-01", NA_complex_, date_lst)
if_else(date_lst > "2020-01-01", as_date(""), date_lst)
```