---
title: "dplyr::recode()"
author: "Yang Zhu"
date: "9/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
```

`dplyr::recode()` can be used along with `dplyr::mutate()` for replacing certain values in a column. 
Sometimes it could be a good alternative for `if_else()`

recode(.x, ..., .default = NULL, .missing = NULL)

argument `.x` is a vector to modify

argument `.default` states the replacement for all values not otherwise matched. 
Please note if replacement is `NA`, it should be something like `NA_character` or `NA_integer`

argument `.missing` states the replacement for missing values

## Replace character object
```{r}
# .default not specified
iris %>% 
  mutate(Species = recode(Species, setosa = "Type 1")) %>% 
  sample_n(size = 10)
# .default has been specified
iris %>% 
  mutate(Species = recode(Species, setosa = "Type 1", .default = NA_character_)) %>% 
  sample_n(size = 10)
```

## Replace numeric object
```{r}
# .default not specified
iris %>% 
  mutate(Sepal.Length = recode(Sepal.Length, `5.5` = 100)) %>% 
  filter(Sepal.Length == 100)
# .default has been specified
iris %>% 
  mutate(Sepal.Length = recode(Sepal.Length, `5.5` = 100, .default = NA_real_)) %>% 
  filter(Sepal.Length == 100 | Sepal.Width < 2.5) %>% 
  sample_n(size = 10)
```