---
title: "dplyr::across()"
author: "Yang Zhu"
date: "9/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
```

`across()` function applies the same transformation to multiple columns, it can be used along with `summarise()` and `mutate()` functions. Similarly, `c_across()` works with `rowwise()` to make row-wise agrregations.

## Usage

across(.cols = everything(), .fns = NULL, ..., .names = NULL) \
c_across(cols = everything())

Note: `.fns` argument can take a function like `mean()`, a purrr-styple lambda \
e.g. `~ mean(.x, na.rm = TRUE)` or a list of functions/lambdas \
e.g. `list(mean = mean, n_miss = ~ sum(is.na(.x))`

## Using function directly

```{r}
iris %>%
  group_by(Species) %>%
  summarise(across(starts_with("Sepal"), mean)) %>% 
  ungroup()

iris %>% 
  mutate(across(starts_with("Sepal"), as.character)) %>%
  glimpse()
```

## Using purrr style

```{r}
iris %>%
  group_by(Species) %>%
  summarise(across(starts_with("Sepal"), ~max(.x, na.rm = TRUE))) %>% 
  ungroup()
```

## Using list of functions

```{r}
iris %>%
  group_by(Species) %>%
  summarise(across(ends_with("Length"), list(mean = mean, sd = sd), .names = "{col}.{fn}")) %>% 
  ungroup()
```

## rowwsie summation

Doing operation on each row

```{r}
iris %>%
  rowwise() %>%
  mutate(
    sum = sum(c_across(Sepal.Length:Petal.Width)),
    sd = sd(c_across(Sepal.Length:Petal.Width))
 )
```