---
title: "dplyr::case_when()"
author: "Yang Zhu"
date: "9/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
```

`case_when()` function allows you to vectorise multiple if_else() statements. If no cases match, NA is returned. \
The difference between `case_when()` and `switch()` is that, usually `case_when()` is used for data manipulation purpose along with `mutate()`, the conditions can be complicated. However, `switch()` is more widely used in function writing, one kind of argument input leads to a completely different data processing method than others.

## Use case_when on data manipulation

```{r}
# get data summary of iris
summary(iris)
# classify iris size by Sepal, Petal length and width
iris_new <- iris %>% 
  mutate(size = case_when(
    Sepal.Length <= 5.1 & Sepal.Width <= 2.8 | Petal.Length <= 1.6 & Petal.Width <= 0.3 ~ "small",
    Sepal.Length >= 6.4 & Sepal.Width >= 3.3 | Petal.Length >= 5.1 & Petal.Width >= 1.8 ~ "large",
    TRUE ~ "medium"
  )) 

iris_new %>% 
  count(size)
# seeing how classification works
table(iris_new$Species, iris_new$size)
```

\newline
We can see most of setosa has small size and most of virginica has large size.



